Site definition possibilites: Features excavated in a transect

### 2018-03-26 analysis/paper/paper.Rmd
added brackets for citations

Because of the test trenching I assume that all features and sites along the transect have been found, even though there are gaps between the actual excavation areas. Machine trenching has been shown to be the most effective way to find sites [@hey_2006]. Admittedly even then not all sites are discovered. Especially neolithic ones have the least probability to turn up, which may be one reason for the small number of middle neolithic features in the transect.

@@

In the American processual discourse a site has been a spatial relationship between archaeological finds [@binford_1992]. Post-processually it has been argued, that it is nothing but a modern analytical category, without any archaeological meaning by itself [@dunnell_1992].

Since the middle of the 1970s there has been the discussion of a non- or off-site archaeology. The terms were introduced by Thomas [-@thomas_1975] and Foley [-@foley_1981] and describe the analysis of finds inbetween sites. The archaeological record was now understood as a continual distribution in space, with spots of higher densities, the so-called "sites", and areas of lower density, which are worthwhile to research as well. With this view on the archaeological record the discussion about the identification of the spatial boundaries of a site arose anew.

As mentioned before, in German archaeology the discussion was not so much focused on "sites" in general but more specifically on the detection of "settlements". 
Next to the problem of qualitatively deciding "What kind of finds or features do I need to call this a settlement", one problem of site detection by survey that has been discussed is the question of how far apart two "sites" are supposed to be. So if I find two sherds of the same time period 50 m apart - are these two sites or belong the sherds together? In every larger settlement or landscape archaeological study this topic is being discussed and in the end a decision felled, which the authors agree, is kind of arbitrary and should actually rely on empiric studies [@malmer_1962; @mischka_2007].

@@

I draw upon the thoughts of Plog et al to define boundaries. They say, sites are *„a discrete and potentially interpretable locus of cultural materials. By discrete, we mean spatially bounded with those boundaries marked by at least relative changes in artifact densities. By interpretable we mean that materials of sufficiently great quality and quantity are present for at least attempting and usually sustaining inferences about the behavior occurring at the locus. By cultural materials we mean artifacts, ecofacts, and features“*. [@plog_1978]

@@

Formal cluster detection analyses as Ripley's K, the F- or G-function [@nakoinz_2016] would rely heavily on edge correction to compensate for the slimness of my transect. I tested them and depending on the different correction algorithms you receive quite different curves. 
Instead I'm not correcting for the edge at all, because admittedly, a transect is all edge. Fully knowing that I only have a very narrow glimpse into the archaeological record, I rely on the length of the transect to give my results credence.

@@

The **Early Bronze Age or Unetice Culture** has only 25 features in this analysis, which are all found in the Western part of the transect. There are five houses and it is notable, that the distances the houses have to each other dominate the distribution. The turning point of the cumulative analysis about 220 m, which correlates quite well with the sizes two known Unetice ring ditches in Schloßvippach, Thuringia [@walter_2007] and  Zwenkau-Eytrah, Saxony [@huth_1998]. Ditches are often interpreted as settlement boundaries.
The distance between settlements which I propose here has an equivalent (430 m) at the excavation of the same street B6n further west, between Benzigerode and Heimburg, where two early bronze age settlements are cut.

@@

As you can see, the **late bronze age** features are those that cluster most densely. The minimal settlement size of about 65 m diameter is comparable to individual late bronze age / early iron age farmsteads that have been recorded at Zwenkau, Saxony [@huth_1998]. 
The cumulative distribution has an insidious slow change of slope, which makes it difficult to determine a turning point. I think this as a sign of not being able to differentiate between different settlements well. This might be due to alternating settlements, meaning people rotated their settlement places faster than we can track archaeologically and several settling events overlap each other. For the settlement in Zwenkau this has been assumed. There the single farmsteads lie more closely together than the miminum calculated by the kernel density estimation.

>>>>

Because of the test trenching I assume that all features and sites along the transect have been found, even though there are gaps between the actual excavation areas. Machine trenching has been shown to be the most effective way to find sites ([@hey_2006]). Admittedly even then not all sites are discovered. Especially neolithic ones have the least probability to turn up, which may be one reason for the small number of middle neolithic features in the transect.

@@

In the American processual discourse a site has been a spatial relationship between archaeological finds ([@binford_1992]). Post-processually it has been argued, that it is nothing but a modern analytical category, without any archaeological meaning by itself ([@dunnell_1992]).

Since the middle of the 1970s there has been the discussion of a non- or off-site archaeology. The terms were introduced by Thomas ([-@thomas_1975]) and Foley ([-@foley_1981]) and describe the analysis of finds inbetween sites. The archaeological record was now understood as a continual distribution in space, with spots of higher densities, the so-called "sites", and areas of lower density, which are worthwhile to research as well. With this view on the archaeological record the discussion about the identification of the spatial boundaries of a site arose anew.

As mentioned before, in German archaeology the discussion was not so much focused on "sites" in general but more specifically on the detection of "settlements". 
Next to the problem of qualitatively deciding "What kind of finds or features do I need to call this a settlement", one problem of site detection by survey that has been discussed is the question of how far apart two "sites" are supposed to be. So if I find two sherds of the same time period 50 m apart - are these two sites or belong the sherds together? In every larger settlement or landscape archaeological study this topic is being discussed and in the end a decision felled, which the authors agree, is kind of arbitrary and should actually rely on empiric studies ([@malmer_1962; @mischka_2007]).

@@

I draw upon the thoughts of Plog et al to define boundaries. They say, sites are *„a discrete and potentially interpretable locus of cultural materials. By discrete, we mean spatially bounded with those boundaries marked by at least relative changes in artifact densities. By interpretable we mean that materials of sufficiently great quality and quantity are present for at least attempting and usually sustaining inferences about the behavior occurring at the locus. By cultural materials we mean artifacts, ecofacts, and features“*. ([@plog_1978])

@@

Formal cluster detection analyses as Ripley's K, the F- or G-function ([@nakoinz_2016]) would rely heavily on edge correction to compensate for the slimness of my transect. I tested them and depending on the different correction algorithms you receive quite different curves. 
Instead I'm not correcting for the edge at all, because admittedly, a transect is all edge. Fully knowing that I only have a very narrow glimpse into the archaeological record, I rely on the length of the transect to give my results credence.

@@

The **Early Bronze Age or Unetice Culture** has only 25 features in this analysis, which are all found in the Western part of the transect. There are five houses and it is notable, that the distances the houses have to each other dominate the distribution. The turning point of the cumulative analysis about 220 m, which correlates quite well with the sizes two known Unetice ring ditches in Schloßvippach, Thuringia ([@walter_2007]) and  Zwenkau-Eytrah, Saxony ([@huth_1998]). Ditches are often interpreted as settlement boundaries.
The distance between settlements which I propose here has an equivalent (430 m) at the excavation of the same street B6n further west, between Benzigerode and Heimburg, where two early bronze age settlements are cut.

@@

As you can see, the **late bronze age** features are those that cluster most densely. The minimal settlement size of about 65 m diameter is comparable to individual late bronze age / early iron age farmsteads that have been recorded at Zwenkau, Saxony ([@huth_1998]). 
The cumulative distribution has an insidious slow change of slope, which makes it difficult to determine a turning point. I think this as a sign of not being able to differentiate between different settlements well. This might be due to alternating settlements, meaning people rotated their settlement places faster than we can track archaeologically and several settling events overlap each other. For the settlement in Zwenkau this has been assumed. There the single farmsteads lie more closely together than the miminum calculated by the kernel density estimation.
